RegressionPvNP

### 2017-05-24 Patient Report.Rmd
…

dataset = 'Training'
#dataset = 'Test'
patient.info = read.patient.info(patient.file, set=dataset)
head(patient.info)

@@

## Thresholds
min.probes=67 # equiv to 1MB
gamma2 = 250
sd.threshold = 0.08  # cov by numbers of probes

gain.threshold = 1.1
loss.threshold = 0.9

cachefile = paste(dataset, '_patients.Rdata', sep='')
if ( file.exists(cachefile) ) {
  load(cachefile, verbose=T)
} else {
  patient.data = list(); lowprobes = c()

@@

      patient = arrange(subset(patient.info, Patient == patient.name),  Endoscopy.Year, Pathology, Samplename)
    }
  }
save(patient.data, file=paste(dataset, '_patients.Rdata', sep=''))
}
print(paste("Missing patients:", paste(setdiff(sum.patient.data$Patient, names(patient.data)), collapse=', ')))

@@

gene.p = gene.p[,(1:4)]
m = melt(gene.p, id.vars=c('Gene.Symbols', 'CN'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))

#m$Gene.Symbols = factor(m$Gene.Symbols, levels=unique(m$Gene.Symbols[c(nrow(m):1)]) ) # order
m$CN = factor(m$CN)

m$brks <- cut(m$value, include.lowest=T, right=F,
                   breaks=c(0, 1, 3, 5, 8, 10),
                   labels=c('0','1-3', '4-5', '6-8','9-10'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))

@@

ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
  labs(title="All available COSMIC genes mapped", x='', y='')

@@

```{r echo=F, warning=F, message=F, fig.height=12, fig.width=8}

m = melt(gene.p[1:50,], id.vars=c('Gene.Symbols', 'CN'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))

@@

ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
  labs(title="Top 50 genes by number of patients", x='', y='')
```

>>>>

dataset = 'Training'
patient.info = read.patient.info(patient.file, set=dataset)
head(patient.info)

@@

## Thresholds
gamma2 = 250
min.probes=67 # equiv to 1MB
sd.threshold = 0.08  # cov by numbers of probes
gain.threshold = 1.1; loss.threshold = 0.9

cachefile = paste(data, '/Analysis/', dataset, '_patients.Rdata', sep='')
if ( file.exists(cachefile) ) {
  load(cachefile, verbose=T)
} else {
  patient.data = list(); lowprobes = c()

@@

      patient = arrange(subset(patient.info, Patient == patient.name),  Endoscopy.Year, Pathology, Samplename)
    }
  }
save(patient.data, file=cachefile)
}
print(paste("Missing patients:", paste(setdiff(sum.patient.data$Patient, names(patient.data)), collapse=', ')))

@@

gene.p = gene.p[,(1:4)]
m = melt(gene.p, id.vars=c('Gene.Symbols', 'CN'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))
m$CN = factor(m$CN)

m$brks <- cut(m$value, include.lowest=T, right=F,ordered_result=T,
                   breaks=c(0, 1, 3, 5, 8, 10),
                   labels=c('0','1-3', '4-5', '6-8','9-10'))

## TODO: Tile the variables so that you can see the gains/losses per progressor/non
subset(m, Gene.Symbols %in% c('CDKN2A', 'TBL1XR1'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, brks))

@@

ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile() + facet_grid(.~variable) + 
    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", x='', y='')

@@

```{r echo=F, warning=F, message=F, fig.height=12, fig.width=8}
gc = as.data.frame(gene.p %>% group_by(Gene.Symbols) %>% summarise(total = sum(Progressors) + sum(Non.Progressors)))
gc = transform(gc, Gene.Symbols = reorder(Gene.Symbols, total))
gc = arrange(gc, -total)

m = melt(subset(gene.p, Gene.Symbols %in% gc[1:50, 'Gene.Symbols']), id.vars=c('Gene.Symbols', 'CN'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))

@@

ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile(color='white') + facet_grid(.~variable) + 
    scale_fill_manual(values=myP, 'Patient Counts') + 
    labs(title="Top 50 genes by number of patients", x='', y='')

#ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
#  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
#  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
#  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
#  labs(title="Top 50 genes by number of patients", x='', y='')
```

### 2017-05-24 RegressionPvNP.Rmd
…

```{r tile, echo=F, message=F, warning=F, include=F}
#tile.w=1e07
tile.w=5e06
tile.files = grep('tile_.*5e\\+06', analysis.files, value=T)
load(tile.files[1], verbose=T)
dim(mergedDf)

cache.dir = paste(data, 'Analysis/5e06', sep='/')
if (!dir.exists(cache.dir)) dir.create(cache.dir)

@@

```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
coefs = list(); plots = list(); performance.at.1se = list()
folds = 10; splits = 5
alpha.values = c(0,0.5,0.7,0.8,0.9,1)

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=1000, family='binomial') # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

@@

```{r xvalpt2, message=F, warning=F, echo=T}

performance.at.1se = do.call(rbind, performance.at.1se)
performance.at.1se$alpha = rownames(performance.at.1se)

@@

ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
       title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

@@

@@

```{r, warning=F, message=F}
#topF = do.call(rbind, strsplit( names(which((coef.stable[['1']]) >= 0.75)), ':|-'))
#colnames(topF) = c('chr','start','end')
#topF = apply(topF, 2, as.numeric)

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  # No HGD/IMC
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
  for (a in alpha.values) {
    # all patients
    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000)

@@

  for (a in alpha.values) {
    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients

@@

```{r noLGD2, echo=T, message=F, warning=F}
performance.at.1se = do.call(rbind, performance.at.1se)
performance.at.1se$alpha = rownames(performance.at.1se)

@@

  few.plots = list()
  for (a in alpha.values) {
    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
    cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitF, minR=0.1)
    #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))

@@

  many.plots = list()
  for (a in alpha.values) {
    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients

@@

```{r leaveoneout, echo=T, message=F, warning=T}

info = do.call(rbind, lapply(patient.data, function(df) df$info))
pg.samp = lapply(patient.data, function(pt) {
  info = pt$info
  print(unique(info$Patient))
  info$SampleSD = NA
  info$SampleMEAN = NA
  if (length(info$Samplename) > 1) {
    info$SampleSD = apply(pt$seg.vals[,info$Samplename], 2, sd)
    info$SampleMEAN =  apply(pt$seg.vals[,info$Samplename], 2, mean)
  }
  info$Prediction = NA
  info$Prediction.Dev.Resid = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

@@

    # Fit generated on all samples, including HGD
    a = 1
    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=1000) # all patients
    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=fitLOO$lambda, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
    #plots[[pt]] = cv$plot + ggtitle(pt)

@@

adj.pred = as.data.frame(all.predictions %>% rowwise() %>% summarise( 
  'pred'=round(correctly.predicted/(sum(correctly.predicted, total.samples )),2),
  'total'=round(total.samples/(sum(correctly.predicted, total.samples )),2),
  'Status'=Status))
adj.pred$Patient = rownames(all.predictions)

adj.pred = transform(adj.pred, Patient=reorder(Patient, -total) )

m = melt(adj.pred, id.vars=c('Status', 'Patient'))
m = arrange(m, Status, variable, -value)
m$Status = as.factor(m$Status)

grid.arrange(
  ggplot(subset(m, Status == 'P') , aes(x=Patient, y=value, fill=variable)) + geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() + labs(title='Progressors',x='', y='ratio'),
  ggplot(subset(m, Status == 'NP') , aes(x=Patient, y=value, fill=variable)) + geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() + labs(title='Non-Progressors',x='', y='ratio')
)

pg = subset(adj.pred, Status == 'P')
np = subset(adj.pred, Status == 'NP')

@@

When predicting each sample individually `r round(nrow(subset(pg, pred == 0.5))/nrow(pg), 2)*100`% patients were correctly predicted in all samples. `r round(nrow(subset(pg, pred < 0.5 & pred > 0))/nrow(pg),2)*100`% of patients were correctly predicted in some samples

`r nrow(subset(pg, pred == 0))` patients had no correctly predicted samples, however, both had only two samples:
```{r, echo=F, warning=F, message=F}
x = do.call(rbind, lapply(pg.samp[subset(pg, pred == 0)$Patient], function(df) df[,c('Status','Patient','Path.ID','Endoscopy.Year', 'Pathology')] ))
row.names(x) = NULL
pander(x)
```

@@

In non-progressors `rnrow(subset(np, pred == 0.5))` patients were predicted to be non-progressors in _all_ of their samples, ranging from `r paste(range(all.predictions[subset(np, pred == 0.5)$Patient, 'total.samples']), collapse='-')`. `r nrow(subset(np, pred >= 0.3))` patients were predicted correctly in 70% or more of their samples. Only `r nrow(subset(np, pred < 0.3))` patients had fewer than 70% of their samples predicted correctly, and both of these have consistently been predicting _progression_ across multiple endoscopies.

`r pander(all.predictions[subset(np, pred < 0.3)$Patient,], caption='Non-progressors that have consistently predicted progression')`

@@

## What clinicicians really want

Backwards from the final sample.

@@

back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
  pt = ddply(pt, .(PID, Endoscopy.Year), summarise, 'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
        'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0 )
  arrange(pt, -Endoscopy.Year, PID)
})

endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples','final.endo')], i)
  }))
  df$Patient = rownames(df)
  return(df)
}))

@@

endos$pred.ratio = round(endos$pred/endos$samples, 2)
endos$i = as.factor(endos$i)

endos = arrange(endos, Patient)
endos = endos %>% group_by(Patient) %>% mutate('n'=length(Patient))
endos = transform(endos, Patient=reorder(Patient, -n) ) 
endos$brks <- cut(endos$pred.ratio, include.lowest=T,
                   breaks=c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1), 
                   labels=c('0', '0.1-.25', '.25-.5', '.5-.75','.75-0.9', '1'))

## TODO This doesn't seem like the most useful way to display this idea
m = melt(endos[,c('Patient','pred.ratio', 'i','n', 'brks')], id.vars = c('Patient', 'i', 'n', 'brks'))

ggplot(m, aes(Patient, i, fill=brks)) + geom_bar(color='grey', stat='identity', position='dodge') + 
  scale_fill_brewer(type='div', palette = "Set2", "Predicted Ratio") + labs(x='', y='Endoscopy from final') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

@@

  r.coefs = list(); r.plots = list()
  folds = 10; splits=5
  for (a in c(0,0.5,1)) {
    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=1000, family='binomial') # all patients

>>>>

```{r tile, echo=F, message=F, warning=F, include=F}
tile.w=1e07
#tile.w=5e06
tile.files = grep('tile_.*1e\\+07', analysis.files, value=T)
#tile.files = grep('tile_.*5e\\+06', analysis.files, value=T)
load(tile.files[1], verbose=T)
dim(mergedDf)

#cache.dir = paste(data, 'Analysis/5e06', sep='/')
cache.dir = paste(data, 'Analysis/1e07', sep='/')
if (!dir.exists(cache.dir)) dir.create(cache.dir)

@@

```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
nl = 3000

coefs = list(); plots = list(); performance.at.1se = list()
folds = 10; splits = 5
alpha.values = c(0,0.5,0.7,0.8,0.9,1)

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial') # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

@@

```{r xvalpt2, message=F, warning=F, echo=T}
performance.at.1se = do.call(rbind, performance.at.1se)
performance.at.1se$alpha = rownames(performance.at.1se)

@@

ggplot(performance.at.1se, aes(alpha, mean)) + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), color='grey') + geom_point() + ylim(0.5,0.8) +
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
       title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

@@

### Features in samples

```{r echo=F, warning=F, message=F}

length(which( coef.stable[['1']] >= 0.5) )

regions = rownames(coefs[['1']])

dysplasia.df[ labels == 1 ,regions]

dysplasia.df[ labels == 0 ,regions]

@@

```{r, warning=F, message=F, eval=F}
#topF = do.call(rbind, strsplit( names(which((coef.stable[['1']]) >= 0.75)), ':|-'))
#colnames(topF) = c('chr','start','end')
#topF = apply(topF, 2, as.numeric)

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  # No HGD/IMC
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
  for (a in alpha.values) {
    # all patients
    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl)

@@

  for (a in alpha.values) {
    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl) # all patients

@@

```{r noLGD2, echo=T, message=F, warning=F}
if (class(performance.at.1se) == 'list')
  performance.at.1se = do.call(rbind, performance.at.1se)

performance.at.1se$alpha = rownames(performance.at.1se)

@@

  few.plots = list()
  for (a in alpha.values) {
    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial') # all patients
    cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitF, minR=0.1)
    #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))

@@

  many.plots = list()
  for (a in alpha.values) {
    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial') # all patients

@@

```{r leaveoneout, echo=T, message=F, warning=F}

info = do.call(rbind, lapply(patient.data, function(df) df$info))
pg.samp = lapply(patient.data, function(pt) {
  info = pt$info
  info$SampleSD = NA
  info$SampleMEAN = NA
  if (length(info$Samplename) > 1) {
    info$SampleSD = apply(pt$seg.vals[,info$Samplename], 2, sd)
    info$SampleMEAN =  apply(pt$seg.vals[,info$Samplename], 2, mean)
  }
  info$Prediction = NA
  info$Prediction.Dev.Resid = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

@@

    # Fit generated on all samples, including HGD
    a = 1
    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl) # all patients
    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=fitLOO$lambda, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
    #plots[[pt]] = cv$plot + ggtitle(pt)

@@

all.predictions$pred.ratio = with(all.predictions, correctly.predicted/total.samples)
all.predictions$Patient = rownames(all.predictions)
all.predictions = transform(all.predictions, Patient=reorder(Patient, pred.ratio) )

m = melt(all.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))

grid.arrange(
  ggplot() + geom_bar(data=subset(m, Status == 'P'), aes(Patient, value), fill='darkblue', color='grey', stat='identity') +  
    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
    labs(title='Progressors', x='', y='Ratio predicted samples'),
  ggplot() + geom_bar(data=subset(m, Status == 'NP'), aes(Patient, value), fill='darkgreen', color='grey', stat='identity') + 
    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
    labs(title='Non Progressors', x='', y='Ratio predicted samples'),
  ncol=2)

pg = subset(all.predictions, Status == 'P')
np = subset(all.predictions, Status == 'NP')

@@

When predicting each sample individually `r round(nrow(subset(pg, pred.ratio == 1))/nrow(pg))*100`% patients were correctly predicted in all samples. `r round(nrow(subset(pg, pred.ratio < 1 & pred.ratio > 0))/nrow(pg),2)*100`% of patients were correctly predicted in some samples

`r nrow(subset(pg, pred.ratio == 0))` patients had no correctly predicted samples.

@@

In non-progressors `r round(nrow(subset(np, pred.ratio == 1))/nrow(np), 2)*100`% of patients were predicted to be non-progressors in _all_ of their samples. `r nrow(subset(np, pred.ratio >= 0.7))` patients were predicted correctly in 70% or more of their samples. Only `r nrow(subset(np, pred.ratio < 0.7 & pred.ratio > 0))` patients had fewer than 70% of their samples predicted correctly, and both of these have consistently been predicting _progression_ across multiple endoscopies.

@@

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

@@

back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
                  'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
                  'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12 )
  arrange(pt, -Endoscopy.Year, PID)
})

endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples','final.endo', 'months.before.final')], i)
  }))
  df$Patient = rownames(df)
  return(df)
}))

@@

endos$pred.ratio = round(endos$pred/endos$samples, 2)
#endos$i = as.factor(endos$i)

#endos$brks <- cut(endos$pred.ratio, include.lowest=T,
#                   breaks=c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1), 
#                   labels=c('0', '0.1-.25', '.25-.5', '.5-.75','.75-0.9', '1'))

endos = transform(endos, Patient=reorder(Patient, -months.before.final) ) 
endos$months.before.final = endos$months.before.final*-1

plot.endo<-function(df, minM=NULL, maxM=NULL) {
  if (is.null(minM)) minM = min(df$months.before.final)
  if (is.null(maxM)) maxM = max(df$months.before.final)

  df$i = as.factor(df$i)
  ggplot(df, aes(months.before.final, pred.ratio, group=1,color=Patient)) + facet_grid(Patient ~., scales='fixed') + scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
    geom_line() + geom_point() + labs(x='Months before diagnosis', y='Sample prediction ratio') + theme(legend.position='none') 
}

me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i))

grid.arrange(
  plot.endo( subset(endos, Patient %in% subset(me, max.endo >=5)$Patient), min(endos$months.before.final), max(endos$months.before.final) ),
  plot.endo( subset(endos, Patient %in% subset(me, max.endo < 5 & max.endo > 2)$Patient),min(endos$months.before.final), max(endos$months.before.final)  ),
  ncol=2)

@@

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
plot.endo( subset(endos, Patient %in% subset(me, max.endo <= 2 & max.endo > 0)$Patient), min(endos$months.before.final), max(endos$months.before.final)  )
```

```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4}
plot.endo( subset(endos, Patient %in% subset(me, max.endo == 0)$Patient),min(endos$months.before.final), max(endos$months.before.final) )
```

### How many samples?  TBD

Clinicians asked how many samples they needed to take.

@@

  r.coefs = list(); r.plots = list()
  folds = 10; splits=5
  for (a in c(0,0.5,1)) {
    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=nl, family='binomial') # all patients
